Agreement Attraction in Turkish: The Case of Genitive Attractors

### 2022-02-10 paper/lcn_revision/manuscript-knitr.Rnw
lcn-revision

\documentclass[apacite,linguex]{glossa}

\let\B\relax
\let\T\relax
\usepackage{xyling} 
\usepackage[linguistics]{forest}

% \pdf* commands provide metadata for the PDF output. ASCII characters only!
\pdfauthor{Utku Turk and Pavel Logacev}
\pdftitle{Agreement Attraction in Turkish}
\pdfkeywords{agreement attraction, syncretism, Turkish, bayesian, replication}

\title[Agreement Attraction in Turkish]{Agreement Attraction in Turkish: The Case of Genitive Attractors\\ \bigskip \large Word count: 3303}

\newcommand{\firstauthor}{{Utku T\"urk}}
\newcommand{\firstauthorshort}{{T\"urk}}
\newcommand{\secondauthor}{{Pavel Loga\v{c}ev}}
\newcommand{\secondauthorshort}{{Loga\v{c}ev}}

\author[\firstauthorshort \& \secondauthorshort]% short form of the author names for the running header. If no short author is given, no authors print in the headers.
{%as many authors as you like, each separated by \AND.
  \spauthor{\firstauthor\\ 
  \institute{Bo\u{g}azi\c{c}i University}\\
  \small{
    \texttt{utku.turk@boun.edu.tr}}
  }
  \AND
  \spauthor{\secondauthor\\ 
  \institute{Bo\u{g}azi\c{c}i University}\\
  \small{
    \texttt{pavel.logacev@boun.edu.tr}}
  }%
}

@@

\SweaveOpts{concordance=TRUE}

\sffamily
\maketitle

\begin{abstract}

Previous studies have shown that speakers may find sentences violating subject-verb agreement grammatical when the sentence contains a feature-matching noun phrase. This so-called agreement attraction effect has also been found in genitive possessive structures such as \textit{'the teacher's brother'} in Turkish (Lago et al., 2019), which is in contrast with its absence in similar constructions in English (Nicol et al., 2016). This discrepancy has been hypothesized to be a result of the association between genitive case marking and subjecthood in Turkish, but not in English. In the present research, we test an alternative explanation in which Turkish number agreement attraction effects are due to a potential confound in Lago et al.'s experiment, as a result of which subject head nouns were locally ambiguous between the possessive and the accusative case. We hypothesized that this ambiguity may have inhibited the availability of the head noun as an agreement controller as the accusative is a non-subject case in Turkish. To test this hypothesis, we conducted a speeded acceptability judgment experiment and our results suggest that case-ambiguity does not play a role in agreement attraction, and thus lends credibility to the claim that genitive noun phrases may function as attractors in Turkish due to the association between genitive case and subjecthood. %exclude

@@

\begin{keywords}
agreement attraction; syncretism; Turkish; bayesian; replication
\end{keywords}

\rmfamily

@@

In this paper, we test an alternative explanation of \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) findings which is related to an instance of local ambiguity in the experimental sentences in their experiment, as a result of which all subject head nouns were ambiguous between possessive and accusative case.

@@

\ex. \label{ex:possgenex}
\gll $[$$[$teknisyen-in$]$ e\u{g}itmen-i$]$ \\
    technician-\textsc{gen} instructor-\textsc{poss} \\
\glt `the instructor of the technician'

\citet{LagoEtAl:2019} demonstrated an agreement attraction effect in Turkish genitive-possessive constructions in a speeded acceptability judgment study with sentences like \ref{ex:lago}, in which the number of the attractor and the verb was manipulated. The resulting $2$x$2$ design, indicated by brackets and slashes in \ref{ex:lago}, consisted of two grammatical conditions, in which the verb agreed with the singular subject head noun, and two ungrammatical conditions, in which the verb carried plural agreement, and thus did not agree with the subject. In ungrammatical sentences, they found a higher percentage of \textit{acceptable} responses when the genitive attractor was plural than when it was singular, indicating agreement attraction. No such effect was found in grammatical sentences.

\ex. \label{ex:lago}
\gll Teknisyen-\{ler/\O\}-in e\u{g}itmen-i ola\u{g}an{\"u}st{\"u} h{\i}zl{\i} ko\c{s}-tu-\{lar/\O\}.\\
technician-\{\textsc{pl}/\textsc{sg}\}-\textsc{gen} instructor-\textsc{poss} extraordinarily fast run-\textsc{pst}-\{\textsc{pl}/\textsc{sg}\}\\
\glt `The technician's/technicians' instructor extraordinarily fast ran\{\textsc{pl}/\textsc{sg}\}.'

\citet{LagoEtAl:2019} hypothesized that these effects originated from how case and number information is encoded and retrieved: According to \citeauthor{LewisVasishth:2005}'s (\citeyear{LewisVasishth:2005}) cue-based retrieval model, phrases are encoded in a content-addressable memory as bundles of features called \textit{chunks} which include information like number, gender, case, and syntactic function \citep[e.g.,][]{SmithVasishth:2020}. Under \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) proposal, participants predict the number of the verb based on the noun phrases they process while reading the subject. In grammatical sentences with singular verb agreement, the number prediction and the verb number match, which causes no processing difficulty. In contrast, when participants fail to find the predicted number morphology on the verb, a memory-retrieval process is initiated. This process activates the search for a chunk matching two cues: the subjecthood feature ([+SUBJECT]) and the plural feature ([+PL]). While neither of the available noun phrases matches this specification in ungrammatical agreement attraction sentences, each of the NPs headed by \textit{technician} and \textit{instructor} matches one of these cues. While this partial match mostly results in participants finding the sentence ungrammatical, they may retrieve the attractor \textit{technicians} on some trials. \citet{LagoEtAl:2019} argue that this erroneous retrieval may be facilitated by the genitive case marking on the attractor, because genitive NPs can function as subjects of embedded clauses in Turkish. Due to the ubiquity of genitive subjects, attractors marked with genitive case are a priori likely to be agreement controllers.

However, a potential problem with the stimuli in the \citet{LagoEtAl:2019} study is that the stems of all head nouns such as \textit{e\u{g}itmen-i} (\textit{`instructor'}) in \ref{ex:lago} were consonant-ending, making their possessive forms ambiguous between possessive and accusative case due to case syncretism between them in vowel-ending stems \citep[pp. 66--67]{GokselKerslake:2005}.
%because the surface form of both, the possessive and accusative case is \textit{-i}. 
%locally
As a result, the head noun (\textit{'instructor'}) in \ref{ex:ambiguous} can be disambiguated towards possessive case as in \ref{ex:ambiguous_possessive}, or towards accusative case as in \ref{ex:ambiguous_accusative}, where the genitive-marked noun functions as an embedded subject, and the matrix subject is omitted due to pro-drop.

\ex.
\label{ex:ambiguous}
  \a. \textsc{Possessive Interpretation} \label{ex:ambiguous_possessive} \\
  \gll Teknisyen-in e\u{g}itmen-i ko\c{s}-tu.\\
       technician-\textsc{gen} instructor-\textsc{poss} run-\textsc{pst}\\
  \glt `The technician's instructor ran.'
  \b. \textsc{Accusative Interpretation} \label{ex:ambiguous_accusative}\\
  \gll Teknisyen-in e\u{g}itmen-i kov-du\u{g}-un-u g\"{o}r-d\"{u}-m.\\
  technician-\textsc{gen} instructor-\textsc{poss} fire-\textsc{nmlz}-\textsc{poss}-\textsc{acc} see-\textsc{pst}-\textsc{1sg}\\
  \glt `I saw the technician firing the fast.'

%{\color{red}
%\ex. %Ambiguous Part 
%\label{ex:ambiguous}\\
%\gll Teknisyen-in e\u{g}itmen-i ola\u{g}an{\"u}st{\"u} h{\i}zl{\i} \ldots \\
%technician-\textsc{gen} instructor-\textsc{poss} extraordinarily fast \ldots \\
%  \a. Simplex Sentence \label{ex:simplexreading} \\
%  \gll \ldots{} ko\c{s}-tu.\\
%  \ldots{} run-\textsc{pst}\\
%  \glt `The technician's instructor extraordinarily fast ran.'
%  \b. Embedded Sentence \label{ex:embeddedreading}\\
%  \gll \ldots{} kov-du\u{g}-un-u g\"{o}r-d\"{u}-m.\\
%  \ldots{} fire-\textsc{nmlz}-\textsc{poss}-\textsc{acc} see-\textsc{pst}-\textsc{1sg}\\
%  \glt `I saw the technician firing the instructor really fast.'
%}

Because accusative NPs cannot function as subjects in Turkish, it is possible that Lago et al.'s finding of agreement attraction effects in sentences like \ref{ex:lago} are not due to the genitive attractors' association with subjecthood, but rather due to the head nouns' \textit{reduced association with subjecthood} due to its ambiguity.
%This ambiguity stems from the fact that the genitive case can serve as an embedded subject marker, and the accusative case can surface as \textit{-i}. 
% For example, on some trials, 
Because sentences like \ref{ex:lago} are locally ambiguous, participants may initially adopt an incorrect analysis of the genitive-possessive structure on some trials, and encode the genitive-marked first noun as the subject of an embedded verb and the ambiguous second noun as an accusative object. 
Under the assumption that remnants of an incorrect analysis affect the parsing process even after after a successful reanalysis \citep{Staub:2007}, the initial association between the genitive noun phrase and subjecthood may lead an agreement attraction effect. 
In contrast, this account predicts that the agreement attraction effect should be either significantly reduced or entirely absent when the head noun is unambiguous, as the function of the genitive noun phrase is disambiguated early, thus preventing potential digging-in effects \citep{Tabor:2004}.  
%
%the head noun will arguably not be treated as a subject until a possible reanalysis after reading the matrix verb. We expect this reduced association to be reflected as an overall decrease in accuracy in grammatical conditions. Upon reading the verb, participants will either find the sentence ungrammatical since they have expected a structure similar to \ref{ex:embeddedreading} or reanalyze the sentence and will judge the sentence according to the structure in \ref{ex:simplexreading}. Since there is no plural marker at the matrix verb, the plural on the attractor will not have an additional effect. However, in ungrammatical conditions in which the matrix verb has an overt plural marker, we expect to have an easier interference of a plural marked Genitive NP which was recently encoded as a subject and an overall decrease in the accuracy. 
%
We tested this hypothesis in a speeded-acceptability experiment with sentences similar to \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}), but with unambiguously marked vowel-ending head nouns. 
%We expect that the agreement attraction effect may arise specifically on trials where participants misparse the two initial NP cases and where the sentence contains a plural attractor with a plural verb. On trials where participants misparse the two initial NP cases, but the sentence does not have a plural attractor or a plural verb, we expect overall higher errors without the specific effect of a plural attractor.
%}

@@

\ex. \label{ex:our_items}
  \a. * \textsc{Plural Attractor, Ungrammatical (Plural Verb)} \label{ex:expitem-plpl}\\
  \gll [Milyoner-ler-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du-lar}.\\
  millionaire-\textsc{pl}-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}-\textsc{pl}\\
  \glt `The millionaires' tailor were fired for no reason at all.'
  \b. \textsc{Plural Attractor, Grammatical (Singular Verb)} \label{ex:expitem-plsg} \\
  \gll [Milyoner-ler-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du}.\\
  millionaire-\textsc{pl}-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}\\
  \glt `The millionaires' tailor was fired for no reason at all.'
  \b. * \textsc{Singular Attractor, Ungrammatical (Plural Verb)} \label{ex:expitem-sgpl}\\
  \gll [Milyoner-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du-lar}.\\
  millionaire-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}-\textsc{pl}\\
  \glt `The millionaire's tailor were fired for no reason at all.'
  \b. \textsc{Singular Attractor Grammatical (Singular Verb)} \label{ex:expitem-sgsg}\\
  \gll [Milyoner-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du}.\\
  millionaire-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}\\
  \glt `The millionaire's tailor was fired for no reason at all.'

@@

We recruited \Sexpr{n_subject} undergraduate students to participate in the experiment in exchange for course credit. All participants were native Turkish speakers, with an average age of \Sexpr{mean_age} (range: \Sexpr{min_age} -- \Sexpr{max_age}). The experiment was carried out following the principles of the Declaration of Helsinki and the regulations concerning research ethics at Bo\u{g}azi\c{c}i University. All participants provided informed consent before their participation and their identity are completely anonimized.

@@

All sentences started with a complex subject NP like \textit{milyonerlerin terzisi} `\textit{the millionaires' tailor},' in which the genitive possessor functioned as the attractor, and the head noun carried an unambiguous possessive case marker. Because the plural marking on nominals is not optional and the head noun was singular, absent of \textit{-lar}, in all conditions, sentences with plural verb agreement were ungrammatical. Moreover, the semantic relationship between the possessor and the head noun was kept as it is in \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) original study and genitive-possessive structures can be paraphrased using \textit{'s} or \textit{of} in English. The distribution of the verb types matched that of the original study, with twenty unergatives, eighteen unaccusatives, and two optionally transitive verbs. Pre-verbal adverbials also consisted of 2-3 words (15 characters on average).

@@

\ex. \label{ex:fillers}
  \a. \textsc{Grammatical Filler (Plural Verb)} \label{ex:gram_filler}\\
  \gll [Sosyolog-un \textbf{\"{o}\u{g}renci-si}] konu\c{s}-unca tutars{\i}zl{\i}k a\c{c}{\i}\u{g}-a \textbf{\c{c}{\i}kar-d{\i}-lar}.\\ 
  sociolog-\textsc{gen}  student-\textsc{poss} speak-\textsc{when} inconsistency  open-\textsc{dat} deduct-\textsc{pst}-\textsc{pl}\\
  \glt `When the student of the sociologist spoke, they revealed an inconsistency.'
  \b. * \textsc{Ungrammatical Filler (Singular Verb)} \label{ex:ung_filler}\\
  \gll [Dans\"{o}z-\"{u}n \textbf{koca-s{\i}}] var-{\i}nca kap{\i} sakince \textbf{a\c{c}-t{\i}}. \\
  dancer-\textsc{gen}  husband-\textsc{poss} arrive-\textsc{when} door slowly  open-\textsc{pst}\\
  \glt Intended:`When the husband of the dancer came, the door opened slowly.'

@@

In order to test whether the morphological ambiguity present in the \citet{LagoEtAl:2019} sentences affected the presence or magnitude of the agreement attraction effect, we analyzed the data from the present experiment compared our results to \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) results, by including \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) data in our Bayesian GLM and using the experiment as an additional factor in the analysis.

Prior to the analysis, we removed the data for all participants who failed to show sufficient sensitivity to the effect of grammaticality in singular attractor conditions, i.e., when no agreement attraction was expected. Specifically, we removed all participants for whom the difference in the percentage of \textit{yes} responses between the grammatical condition \ref{ex:expitem-sgsg} and the ungrammatical condition \ref{ex:expitem-sgpl} fell below the threshold of 0.25 percentage points. We also excluded trials in which the participants missed the response deadline or gave overly fast responses (below 200 ms). As a result, we excluded 10.22\% of the trials from our experiment and 2.38\% of the \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) trials.

We analyzed responses using two Bayesian GLMs assuming a Bernoulli-distributed response with a probit link function.  We used the R packages brms \citep{brms} and rstan \citep{rstan} to fit Bayesian hierarchical models \citep[e.g.,][]{GelmanHill:2007, NicenboimVasishth:2016}. 
We analyzed only experimental sentences and used (i) grammaticality of the sentence, (ii) attractor number, and (iii) presence of morphological ambiguity (i.e., experiment), as well as all their interactions as predictors. We used by-participant and by-item intercepts and slopes for all predictors. 
All factors were sum-coded. We have used standard priors provided by the brms package.

Because the magnitude of the agreement attraction effect can be operationalized either as the interaction between grammaticality and the presence of a plural attractor, or as the effect of a plural attractor in ungrammatical sentences, we used a second model with predictors (ii) and (iii) and their interaction to analyze responses to ungrammatical conditions only.

In the results section, we provide summaries of the coefficient posterior distributions. For our model, we ran \Sexpr{n_chains} chains with \Sexpr{n_warmup} warm-up iterations and \Sexpr{n_iter-n_warmup} sampling iterations. The data for our study, along with our analysis scripts can be found \url{https://osf.io/hegmd/}.

@@

\caption{The average percentage of acceptable responses according to the experimental conditions in our study and \citet{LagoEtAl:2019}. Error bars signal standard errors calculated following \citet{Morey:2008,Cousineau:2005}.}
\label{fig:AverageResponses}
\end{figure}

@@

\begin{figure}[hbt!]
\centering

<<ResponseModel,  echo=FALSE, fig.align='center', height=3, warning=FALSE, dev = cairo_pdf, dpi = 300, message=FALSE>>=

>>>>

% interactapasample.tex
% v1.05 - August 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled

\usepackage{tikz,xcolor,hyperref}

% Make Orcid icon
\definecolor{lime}{HTML}{A6CE39}
\DeclareRobustCommand{\orcidicon}{%
	\begin{tikzpicture}
	\draw[lime, fill=lime] (0,0) 
	circle [radius=0.16] 
	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
	\draw[white, fill=white] (-0.0625,0.095) 
	circle [radius=0.007];
	\end{tikzpicture}
	\hspace{-2mm}
}

\foreach \x in {A, ..., Z}{%
	\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}{\noexpand\orcidicon}}
}

% Define the ORCID iD command for each author separately. Here done for two authors.
\newcommand{\orcidauthorA}{0000-0001-8011-1541}
\newcommand{\orcidauthorB}{0000-0002-2837-2258}

%\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
%\bibliographystyle{unsrtnat}
%\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
%\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty

\usepackage[natbibapa,nodoi]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!
\setlength\bibhang{12pt}% To set the indentation in the list of references using apacite.sty. Commands using natbib.sty MUST be deactivated first!
%\let\cite\shortcite

\renewcommand\bibliographytypesize{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using apacite.sty. Commands using natbib.sty MUST be deactivated first!

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\let\B\relax
\let\T\relax
\usepackage{xyling} 
\usepackage[linguistics]{forest}
\usepackage{linguex}
\renewcommand{\firstrefdash}{}%
\AtBeginDocument{\settowidth{\Exlabelwidth}{(110)}}
\RequirePackage{cgloss} %for adding the language name and source of the example on the first line of glossed examples (requires \gll before the foreign language example and \glt before the translation)

\newcommand{\rev}[1]{{\color{red}#1}}

@@

\articletype{Short Article}% Specify the article type or omit as appropriate

\title{Agreement Attraction in Turkish: The Case of Genitive Attractors}

\author{
\name{Utku T\"{u}rk\textsuperscript{a}\orcidA{}\thanks{CONTACT Utku T\"{u}rk. Email: utkuturk@pm.me} and Pavel Loga\v{c}ev\textsuperscript{a}\orcidB{}}
\affil{\textsuperscript{a}Bo\u{g}azi\c{c}i University, \.{I}stanbul, Turkey}
}

\maketitle

\begin{abstract}
Speakers have been shown to find sentences with erroneous agreement acceptable under certain conditions. This so-called agreement attraction effect has also been found in genitive-possessive structures such as \textit{``the teacher's brother''} in Turkish (Lago et al., 2019), which is in contrast to its absence in similar constructions in English (Nicol et al., 2016). It has been hypothesized that this discrepancy is a result of the association between genitive case marking and subjecthood in Turkish. We test an alternative explanation according to which Lago et al.'s findings are due to a potential confound in their experiment, as all subject head nouns were locally ambiguous between possessive and accusative case. The results of our speeded acceptability judgment experiment suggest that the presence of case ambiguity does not affect agreement attraction, and thus suggests that genitive NPs may function as attractors in Turkish due to the association between genitive case and subjecthood.

@@

\begin{keywords}
Agreement attraction; syncretism; Turkish; \rev{sentence processing; number}
\end{keywords}

@@

In this paper, we test an alternative explanation of \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) findings which is related to an instance of local ambiguity in \rev{their} experimental sentences, as a result of which all subject head nouns were ambiguous between possessive and accusative case.

@@

\ex. \label{ex:possgenex}
\gll $[$$[$teknisyen-in$]$ e\u{g}itmen-i$]$ \\
    technician-\textsc{gen} instructor-\textsc{poss} \\
\glt ``the instructor of the technician''

\citet{LagoEtAl:2019} demonstrated an agreement attraction effect in Turkish genitive-possessive constructions in a speeded acceptability judgment study with sentences like \ref{ex:lago}, in which the number of the attractor and the verb was manipulated. The resulting $2$x$2$ design, indicated by brackets and slashes in \ref{ex:lago}, consisted of two grammatical conditions, in which the verb agreed \rev{in number} with the singular subject head noun, and two ungrammatical conditions, in which the verb carried plural agreement, and thus did not agree with the subject. In ungrammatical sentences, they found a higher percentage of \textit{acceptable} responses when the genitive attractor was plural than when it was singular, indicating agreement attraction. No such effect was found in grammatical sentences.

\ex. \label{ex:lago}
\gll Teknisyen-\{ler/\O\}-in e\u{g}itmen-i ola\u{g}an{\"u}st{\"u} h{\i}zl{\i} ko\c{s}-tu-\{lar/\O\}.\\
technician-\{\textsc{pl}/\textsc{sg}\}-\textsc{gen} instructor-\textsc{poss} extraordinarily fast run-\textsc{pst}-\{\textsc{pl}/\textsc{sg}\}\\
\glt ``The technician's/technicians' instructor \rev{ran\{\textsc{pl}/\textsc{sg}\}} extraordinarily fast.''

\citet{LagoEtAl:2019} hypothesized that these effects originated from how case and number information \rev{are} encoded and retrieved: According to \citeauthor{LewisVasishth:2005}'s (\citeyear{LewisVasishth:2005}) cue-based retrieval model, phrases are encoded in a content-addressable memory as bundles of features called \textit{chunks} which include information like number, gender, case, and syntactic function \citep[e.g.,][]{SmithVasishth:2020}. Under \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) proposal, participants predict the number of the verb based on the noun phrases they process while reading the subject. In grammatical sentences with singular verb agreement, the number prediction and the verb number match, which causes no processing difficulty. In contrast, when participants fail to find the predicted number morphology on the verb, a memory-retrieval process is initiated. This process activates the search for a chunk matching two cues: the subjecthood feature ([+SUBJECT]) and the plural feature ([+PL]). While neither of the available noun phrases matches this specification in ungrammatical agreement attraction sentences, each of the NPs headed by \textit{technician} and \textit{instructor} matches one of these cues. While this partial match mostly results in participants finding the sentence ungrammatical, they may retrieve the attractor \textit{technicians} on some trials. \citet{LagoEtAl:2019} argue that this erroneous retrieval may be facilitated by the genitive case marking on the attractor, because genitive NPs can function as subjects of embedded clauses in Turkish. Due to the ubiquity of genitive subjects, attractors marked with genitive case are a priori likely to be agreement controllers.

However, a potential problem with the stimuli in the \citet{LagoEtAl:2019} study is that the stems of all head nouns such as \textit{e\u{g}itmen-i} (\textit{`instructor'}) in \ref{ex:lago} were consonant-ending, making their possessive forms ambiguous between possessive and accusative case due to case syncretism between them in vowel-ending stems \citep[pp. 66--67]{GokselKerslake:2005}. As a result, the head noun (\textit{'instructor'}) in \ref{ex:ambiguous} can be disambiguated towards possessive case as in \ref{ex:ambiguous_possessive}, or towards accusative case as in \ref{ex:ambiguous_accusative}, where the genitive-marked noun functions as an embedded subject, and the matrix subject is omitted due to pro-drop. 
\rev{We also calculated the relative probability of having an accusative marked noun compared to an possessive marked noun following genitive marking. Data from annotated treebanks from Universal Dependencies v2.9 \citep{TurkEtAl2021,KuzgunEtAl2020,TurkEtAl2019b,SulubacakEtAl2016,Coltekin2015} showed that the relative probability of encountering accusative marking after the genitive-marked noun is \Sexpr{p_acc_given_gen}.}

\ex.
\label{ex:ambiguous}
  \a. \textsc{Possessive Interpretation} \label{ex:ambiguous_possessive} \\
  \gll Teknisyen-in e\u{g}itmen-i ko\c{s}-tu.\\
       technician-\textsc{gen} instructor-\textsc{poss} run-\textsc{pst}\\
  \glt ``The technician's instructor ran.''
  \b. \textsc{Accusative Interpretation} \label{ex:ambiguous_accusative}\\
  \gll Teknisyen-in e\u{g}itmen-i kov-du\u{g}-un-u g\"{o}r-d\"{u}-m.\\
  technician-\textsc{gen} instructor-\textsc{poss} fire-\textsc{nmlz}-\textsc{poss}-\textsc{acc} see-\textsc{pst}-\textsc{1sg}\\
  \glt ``I saw the technician firing the \rev{instructor}.''

Because accusative NPs cannot function as subjects in Turkish, it is possible that Lago et al.'s finding of agreement attraction effects in sentences like \ref{ex:lago} are not due to the genitive attractors' association with subjecthood, but rather due to the head nouns' \textit{reduced association with subjecthood} due to its ambiguity.

Because sentences like \ref{ex:lago} are locally ambiguous, participants may initially adopt an incorrect analysis of the genitive-possessive structure on some trials, and encode the genitive-marked first noun as the subject of an embedded verb and the ambiguous second noun as an accusative object. 
Under the assumption that remnants of an incorrect analysis affect the parsing process even after after a successful reanalysis \citep{Staub:2007}, the initial association between the genitive noun phrase and subjecthood may lead an agreement attraction effect. 
In contrast, this account predicts that the agreement attraction effect should be either significantly reduced or entirely absent when the head noun is unambiguous, as the function of the genitive noun phrase is disambiguated early, thus preventing potential digging-in effects \citep{Tabor:2004}. We tested this hypothesis in a speeded-acceptability experiment with sentences similar to \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}), but with unambiguously marked vowel-ending head nouns.

@@

\ex. \label{ex:our_items}
  \a. * \textsc{Plural Attractor, Ungrammatical (Plural Verb)} \label{ex:expitem-plpl}\\
  \gll [Milyoner-ler-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du-lar}.\\
  millionaire-\textsc{pl}-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}-\textsc{pl}\\
  \glt ``The millionaires' tailor were fired for no reason at all.''
  \b. \textsc{Plural Attractor, Grammatical (Singular Verb)} \label{ex:expitem-plsg} \\
  \gll [Milyoner-ler-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du}.\\
  millionaire-\textsc{pl}-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}\\
  \glt ``The millionaires' tailor was fired for no reason at all.''
  \b. * \textsc{Singular Attractor, Ungrammatical (Plural Verb)} \label{ex:expitem-sgpl}\\
  \gll [Milyoner-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du-lar}.\\
  millionaire-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}-\textsc{pl}\\
  \glt ``The millionaire's tailor were fired for no reason at all.''
  \b. \textsc{Singular Attractor Grammatical (Singular Verb)} \label{ex:expitem-sgsg}\\
  \gll [Milyoner-in \textbf{terzi-si}] tamamen gereksizce \textbf{kov-ul-du}.\\
  millionaire-\textsc{gen} tailor-\textsc{poss} completely without.reason fire-\textsc{pass}-\textsc{pst}\\
  \glt ``The millionaire's tailor was fired for no reason at all.''

@@

We recruited \Sexpr{n_subject} undergraduate students to participate in the experiment in exchange for course credit. All participants were native Turkish speakers, with an average age of \Sexpr{mean_age} (range: \Sexpr{min_age} -- \Sexpr{max_age}). The experiment was carried out following the principles of the Declaration of Helsinki and the regulations concerning research ethics at Bo\u{g}azi\c{c}i University. All participants provided informed consent before their participation and their identity \rev{were} completely anonimised.

@@

All sentences started with a complex subject NP like \textit{milyonerlerin terzisi} ``\textit{the millionaires' tailor},'' in which the genitive possessor functioned as the attractor, and the head noun carried an unambiguous possessive case marker. Because the plural marking on nominals is not optional and the head noun was singular, absent of \textit{-lar}, in all conditions, sentences with plural verb agreement were ungrammatical. Moreover, the semantic relationship between the possessor and the head noun was kept as it is in \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) original study and genitive-possessive structures can be paraphrased using \textit{'s} or \textit{of} in English. The distribution of the verb types matched that of the original study, with twenty unergatives, eighteen unaccusatives, and two optionally transitive verbs. Pre-verbal adverbials also consisted of 2-3 words (15 characters on average).

@@

\ex. \label{ex:fillers}
  \a. \textsc{Grammatical Filler (Plural Verb)} \label{ex:gram_filler}\\
  \gll [Sosyolog-un \textbf{\"{o}\u{g}renci-si}] konu\c{s}-unca tutars{\i}zl{\i}k a\c{c}{\i}\u{g}-a \textbf{\c{c}{\i}kar-d{\i}-lar}.\\ 
  sociolog-\textsc{gen}  student-\textsc{poss} speak-\textsc{when} inconsistency  open-\textsc{dat} deduct-\textsc{pst}-\textsc{pl}\\
  \glt ``When the student of the sociologist spoke, they revealed an inconsistency.''
  \b. * \textsc{Ungrammatical Filler (Singular Verb)} \label{ex:ung_filler}\\
  \gll [Dans\"{o}z-\"{u}n \textbf{koca-s{\i}}] var-{\i}nca kap{\i} sakince \textbf{a\c{c}-t{\i}}. \\
  dancer-\textsc{gen}  husband-\textsc{poss} arrive-\textsc{when} door slowly  open-\textsc{pst}\\
  \glt Intended:``When the husband of the dancer came, the door opened slowly.''

@@

In order to test whether the morphological ambiguity present in the \citet{LagoEtAl:2019} sentences affected the presence or magnitude of the agreement attraction effect, we analyzed the data from the present experiment \rev{and} compared our results to \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) results, by including \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) data in our Bayesian GLM and using the experiment as an additional factor in the analysis.

Prior to the analysis, we removed the data for all participants who failed to show sufficient sensitivity to the effect of grammaticality in singular attractor conditions, i.e., when no agreement attraction was expected. Specifically, we removed all participants for whom the difference in the percentage of \textit{yes} responses between \rev{conditions with singular attractors \ref{ex:expitem-sgsg} and \ref{ex:expitem-sgpl}} fell below the threshold of \Sexpr{accuracy_threshold} percentage points. We also excluded trials in which the participants missed the response deadline. As a result, we excluded \rev{\Sexpr{perc_deletion_exp1}\%} of the trials from our experiment and \Sexpr{perc_deletion_lago}\% of the \citeauthor{LagoEtAl:2019}'s (\citeyear{LagoEtAl:2019}) trials.

We analyzed responses using two Bayesian GLMs assuming a Bernoulli-distributed response with a probit link function.  We used the R packages brms \citep{brms} and rstan \citep{rstan} to fit Bayesian hierarchical models \citep[e.g.,][]{GelmanHill:2007, NicenboimVasishth:2016}. 
We analyzed only experimental sentences and used (i) grammaticality of the sentence, (ii) attractor number, and (iii) presence of morphological ambiguity (i.e., experiment), as well as all their interactions as predictors. We used by-participant and by-item intercepts and slopes for all predictors and \rev{their interaction}. 
All factors were sum-coded. \rev{A Student's \emph{t}-distribution ($\nu$ = 3, $\mu$ = 0, $\sigma$ = 2.5), a normal distribution ($\mu$ = 0, $\sigma$ = 1), a Cauchy distribution (($\mu$ = 0, $\sigma$ = 1)) and a LKJ distribution ($\eta$ = 2) were used for our intercept, slopes of predictors, standard deviations of random effects, correlation coefficients in interaction models respectively as priors.}

Because the magnitude of the agreement attraction effect can be operationalised either as the interaction between grammaticality and the presence of a plural attractor, or as the effect of a plural attractor in ungrammatical sentences, we used a second model with predictors (ii) and (iii) and their interaction to analyze responses to ungrammatical conditions only.

In the results section, we provide summaries of the coefficient posterior distributions. For our model, we ran \Sexpr{n_chains} chains with \Sexpr{n_warmup} warm-up iterations and \Sexpr{n_iter-n_warmup} sampling iterations. The data for our study, along with our analysis scripts can be found at \url{https://osf.io/hegmd/}.

@@

@@

<<ResponseModel, message = FALSE, echo=FALSE,  warning = FALSE, dev='cairo_pdf', fig.height=3, fig.align='center', out.width='\\linewidth', fig.pos='hbt!', fig.cap='Estimates and 95\\% credible intervals for the probit regression coefficients for the model of responses in our experiment and \\citet{LagoEtAl:2019}.'>>=
